---
title: "Team 26 Stat 3355 Project"
author: "Kenan Stredic, Roopali Kallem"
date: "2022-11-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


importing necessary libraries

```{r}

library(ggplot2)

library(gridExtra)

```

reading the csv files into data frames

```{r}

# reading the files into data frames

world_covid <- read.csv( "WHO C-19 cummulative.csv" )
air_traffic <- read.csv( "c-19 impact on air traffic.csv" )
world_vacc <- read.csv( "vaccination-data.csv" )

```

Data Pre-Processing

```{r}

# converting dates in the data frames from character type to date type.

# Factor Date_reported in world_covid
world_covid_Date_reported_factored <- as.factor( world_covid$Date_reported )

# Convert Date_reported from character to POSIXlt in world_covid
world_covid_Date_reported_POSIXlt <- strptime( world_covid_Date_reported_factored, format = "%m/%d/%Y" )

# Convert Date_reported from POSIXlt to Date in world_covid
world_covid$Date_reported <- as.Date( world_covid_Date_reported_POSIXlt, format = "%Y-%m-%d" )

str( world_covid ) # printing to see the data type of each variable

```

```{r}

# Removing un-necessary columns from world_covid

world_covid <- world_covid[ -c( 2, 4 ) ]
str( world_covid ) # printing again to see if correct columns are removed and to confirm data type of each variable


# checking for any possible null values

sum( is.na( world_covid$Date_reported ) )
sum( is.na( world_covid$Country ) ) 
sum( is.na( world_covid$New_cases ) )
sum( is.na( world_covid$Cumulative_cases ) )
sum( is.na( world_covid$New_deaths ) )
sum( is.na( world_covid$Cumulative_deaths ) )

```


```{r}

# converting dates in the data frames from character type to date type.

# Factor Date_reported in air_traffic
air_traffic_Date_factored <- as.factor( air_traffic$Date ) 

# Convert Date from character to POSIXlt in air_traffic
air_traffic_Date_POSIXlt <- strptime( air_traffic_Date_factored, format = "%m/%d/%Y" )

# Convert Date from POSIXlt to Date in air_traffic
air_traffic$Date <- as.Date( air_traffic_Date_POSIXlt, format = "%Y-%m-%d" )

str( air_traffic ) # printing to see the data type of each variable

```


```{r}

# Removing un-necessary columns from air_traffic
air_traffic <- air_traffic[ -c( 1, 3, 4, 6, 9, 11 ) ]
str( air_traffic )  # printing again to see if correct columns are removed and to confirm data type of each variable



# checking for any possible null values

sum( is.na( air_traffic$Date ) )
sum( is.na( air_traffic$Country ) )
sum( is.na( air_traffic$City ) )
sum( is.na( air_traffic$State ) )

```


```{r}

# Factor Date_updated in world_vacc
world_vacc_Date_Updated_factored <- as.factor( world_vacc$DATE_UPDATED )

# Convert Date_updated from character to POSIXlt in world_vacc
world_vacc_Date_Updated_POSIXlt <- strptime( world_vacc_Date_Updated_factored, format = "%m/%d/%Y" )

# Convert Date_updated from POSIXlt to Date in world_vacc
world_vacc$DATE_UPDATED <- as.Date( world_vacc_Date_Updated_POSIXlt, format = "%Y-%m-%d" )



# Factor FIRST_VACCINE_DATE in world_vacc
world_vacc_FIRST_VACCINE_DATE_factored <- as.factor( world_vacc$FIRST_VACCINE_DATE )

# Convert FIRST_VACCINE_DATE from character to POSIXlt in world_vacc
world_vacc_FIRST_VACCINE_DATE_POSIXlt <- strptime( world_vacc_FIRST_VACCINE_DATE_factored, format = "%m/%d/%Y" )

# Convert FIRST_VACCINE_DATE from POSIXlt to Date in world_vacc
world_vacc$FIRST_VACCINE_DATE <- as.Date( world_vacc_FIRST_VACCINE_DATE_POSIXlt, format = "%Y-%m-%d" )

str( world_vacc )  # printing to see the data type of each variable


```


```{r}

# Removing un-necessary columns from world_covid
world_vacc <- world_vacc[ -c( 2:4, 12:16 ) ]
str( world_vacc )  # printing again to see if correct columns are removed and to confirm data type of each variable


# checking for any possible null values

sum( is.na( world_vacc$DATE_UPDATED ) )
sum( is.na( world_vacc$Country ) )
sum( is.na( world_vacc$TOTAL_VACCINATIONS ) )
sum( is.na( world_vacc$PERSONS_VACCINATED_1PLUS_DOSE ) )
sum( is.na( world_vacc$TOTAL_VACCINATIONS_PER100 ) )
sum( is.na( world_vacc$PERSONS_VACCINATED_1PLUS_DOSE_PER100 ) )
sum( is.na( world_vacc$PERSONS_FULLY_VACCINATED ) )
sum( is.na( world_vacc$PERSONS_FULLY_VACCINATED_PER100 ) )

```


```{r}

# replacing the Null values with the mean of the column

world_vacc$TOTAL_VACCINATIONS[ is.na( world_vacc$TOTAL_VACCINATIONS ) ] <- mean( world_vacc$TOTAL_VACCINATIONS, na.rm = TRUE )

world_vacc$PERSONS_VACCINATED_1PLUS_DOSE[ is.na( world_vacc$PERSONS_VACCINATED_1PLUS_DOSE ) ] <- mean( world_vacc$PERSONS_VACCINATED_1PLUS_DOSE, na.rm = TRUE )

world_vacc$TOTAL_VACCINATIONS_PER100[ is.na( world_vacc$TOTAL_VACCINATIONS_PER100 ) ] <- mean( world_vacc$TOTAL_VACCINATIONS_PER100, na.rm = TRUE )

world_vacc$PERSONS_VACCINATED_1PLUS_DOSE_PER100[ is.na( world_vacc$PERSONS_VACCINATED_1PLUS_DOSE_PER100 ) ] <- mean( world_vacc$PERSONS_VACCINATED_1PLUS_DOSE_PER100, na.rm = TRUE )

world_vacc$PERSONS_FULLY_VACCINATED[ is.na( world_vacc$PERSONS_FULLY_VACCINATED ) ] <- mean( world_vacc$PERSONS_FULLY_VACCINATED, na.rm = TRUE )

world_vacc$PERSONS_FULLY_VACCINATED_PER100[ is.na( world_vacc$PERSONS_FULLY_VACCINATED_PER100 ) ] <- mean( world_vacc$PERSONS_FULLY_VACCINATED_PER100, na.rm = TRUE )

# rechecking to see if any more null values

sum( is.na( world_vacc$DATE_UPDATED ) )
sum( is.na( world_vacc$Country ) ) 
sum( is.na( world_vacc$TOTAL_VACCINATIONS ) )
sum( is.na( world_vacc$PERSONS_VACCINATED_1PLUS_DOSE ) )
sum( is.na( world_vacc$TOTAL_VACCINATIONS_PER100 ) )
sum( is.na( world_vacc$PERSONS_VACCINATED_1PLUS_DOSE_PER100 ) )
sum( is.na( world_vacc$PERSONS_FULLY_VACCINATED ) )
sum( is.na( world_vacc$PERSONS_FULLY_VACCINATED_PER100 ) )

```

Data Visualization


Between the two years is there any pattern between the number of cases and the number of deaths 

```{r}
# Between the two years is there any pattern between the number of cases and the number of deaths in USA
# Instead of using just the smooth line graph, we showed the line and point graph next to the smooth line graph 

usa_covid <- world_covid[ which( world_covid$Country == "United States of America" ), ]

usa_deaths1 <- ggplot( data = usa_covid, mapping = aes( x = Date_reported, y = New_deaths ) ) + 
  geom_line() + 
  geom_point() +
  labs( x = "Date Reported", y = "New Deaths", title = "Covid-19 Deaths in the USA" )

usa_deaths2 <- ggplot( data = usa_covid, mapping = aes( x = Date_reported, y = New_deaths ) ) + 
  geom_smooth() + 
  labs( x = "Date Reported", y = "New Deaths", title = "Covid-19 Deaths in the USA" )

usa_cases1 <- ggplot( data = usa_covid, mapping = aes( x = Date_reported, y = New_cases ) ) + 
  geom_line() + 
  geom_point() +
  labs( x = "Date Reported", y = "New Cases", title = "Covid-19 Cases in the USA" )

usa_cases2 <- ggplot( data = usa_covid, mapping = aes( x = Date_reported, y = New_cases ) ) + 
  geom_smooth() +
  labs( x = "Date Reported", y = "New Cases", title = "Covid-19 Cases in the USA" )

grid.arrange( usa_deaths1, usa_deaths2, nrow = 1 )
grid.arrange( usa_cases1, usa_cases2, nrow = 1 )

# Between the two years is there any pattern between the number of cases and the number of deaths in India

india_covid <- world_covid[ which( world_covid$Country == "India" ), ]

india_deaths1 <- ggplot( data = india_covid, mapping = aes( x = Date_reported, y = New_deaths ) ) + 
  geom_line() + 
  geom_point() +
  labs( x = "Date Reported", y = "New Deaths", title = "Covid-19 Deaths in India" )

india_deaths2 <- ggplot( data = india_covid, mapping = aes( x = Date_reported, y = New_deaths ) ) + 
  geom_smooth() + 
  labs( x = "Date Reported", y = "New Deaths", title = "Covid-19 Deaths in India" )

india_cases1 <- ggplot( data = india_covid, mapping = aes( x = Date_reported, y = New_cases ) ) + 
  geom_line() + 
  geom_point() +
  labs( x = "Date Reported", y = "New Cases", title = "Covid-19 Cases in India" )

india_cases2 <- ggplot( data = india_covid, mapping = aes( x = Date_reported, y = New_cases ) ) + 
  geom_smooth() +
  labs( x = "Date Reported", y = "New Cases", title = "Covid-19 Cases in India" )

grid.arrange( india_deaths1, india_deaths2, nrow = 1 )
grid.arrange( india_cases1, india_cases2, nrow = 1 )

# Between the two years is there any pattern between the number of cases and the number of deaths in Germany

germany_covid <- world_covid[ which( world_covid$Country == "Germany" ), ]

germany_deaths1 <- ggplot( data = germany_covid, mapping = aes( x = Date_reported, y = New_deaths ) ) + 
  geom_line() + 
  geom_point() +
  labs( x = "Date Reported", y = "New Deaths", title = "Covid-19 Deaths in Germany" )

germany_deaths2 <- ggplot( data = germany_covid, mapping = aes( x = Date_reported, y = New_deaths ) ) + 
  geom_smooth() + 
  labs( x = "Date Reported", y = "New Deaths", title = "Covid-19 Deaths in Germany" )

germany_cases1 <- ggplot( data = germany_covid, mapping = aes( x = Date_reported, y = New_cases ) ) + 
  geom_line() + 
  geom_point() +
  labs( x = "Date Reported", y = "New Cases", title = "Covid-19 Cases in Germany" )

germany_cases2 <- ggplot( data = germany_covid, mapping = aes( x = Date_reported, y = New_cases ) ) + 
  geom_smooth() +
  labs( x = "Date Reported", y = "New Cases", title = "Covid-19 Cases in Germany" )

grid.arrange( germany_deaths1, germany_deaths2, nrow = 1 )
grid.arrange( germany_cases1, germany_cases2, nrow = 1 )

# Between the two years is there any pattern between the number of cases and the number of deaths in France

france_covid <- world_covid[ which( world_covid$Country == "France" ), ]

france_deaths1 <- ggplot( data = france_covid, mapping = aes( x = Date_reported, y = New_deaths ) ) + 
  geom_line() + 
  geom_point() +
  labs( x = "Date Reported", y = "New Deaths", title = "Covid-19 Deaths in France" )

france_deaths2 <- ggplot( data = france_covid, mapping = aes( x = Date_reported, y = New_deaths ) ) + 
  geom_smooth() + 
  labs( x = "Date Reported", y = "New Deaths", title = "Covid-19 Deaths in France" )

france_cases1 <- ggplot( data = france_covid, mapping = aes( x = Date_reported, y = New_cases ) ) + 
  geom_line() + 
  geom_point() +
  labs( x = "Date Reported", y = "New Cases", title = "Covid-19 Cases in France" )

france_cases2 <- ggplot( data = france_covid, mapping = aes( x = Date_reported, y = New_cases ) ) + 
  geom_smooth() +
  labs( x = "Date Reported", y = "New Cases", title = "Covid-19 Cases in France" )

grid.arrange( france_deaths1, france_deaths2, nrow = 1 )
grid.arrange( france_cases1, france_cases2, nrow = 1 )

# Between the two years is there any pattern between the number of cases and the number of deaths in Brazil

brazil_covid <- world_covid[ which( world_covid$Country == "Brazil" ), ]

brazil_deaths1 <- ggplot( data = brazil_covid, mapping = aes( x = Date_reported, y = New_deaths ) ) + 
  geom_line() + 
  geom_point() +
  labs( x = "Date Reported", y = "New Deaths", title = "Covid-19 Deaths in Brazil" )

brazil_deaths2 <- ggplot( data = brazil_covid, mapping = aes( x = Date_reported, y = New_deaths ) ) + 
  geom_smooth() + 
  labs( x = "Date Reported", y = "New Deaths", title = "Covid-19 Deaths in Brazil" )

brazil_cases1 <- ggplot( data = brazil_covid, mapping = aes( x = Date_reported, y = New_cases ) ) + 
  geom_line() + 
  geom_point() +
  labs( x = "Date Reported", y = "New Cases", title = "Covid-19 Cases in Brazil" )

brazil_cases2 <- ggplot( data = brazil_covid, mapping = aes( x = Date_reported, y = New_cases ) ) + 
  geom_smooth() +
  labs( x = "Date Reported", y = "New Cases", title = "Covid-19 Cases in Brazil" )

grid.arrange( brazil_deaths1, brazil_deaths2, nrow = 1 )
grid.arrange( brazil_cases1, brazil_cases2, nrow = 1 )

```

Did air traffic play a role in the increase of covid cases at the start of the pandemic (2020)?

```{r}

library( dplyr )

usa_air_traffic <- air_traffic[ which( air_traffic$Country == 'United States of America (the)' ), ]

usa_air_traffic <- usa_air_traffic %>% arrange( Date )

usa_air_traffic[, "Cumulative_flights" ] <- cumsum( usa_air_traffic$PercentOfBaseline )


usa_20 <- usa_covid %>% filter(between(Date_reported, as.Date('2020-03-16'), as.Date('2020-12-02')))

# Changed from bar plots to line plots

usa_cases <- ggplot( data = usa_20, aes( x = Date_reported, y = Cumulative_cases ) ) + 
  geom_line( stat = "identity", color = "red" ) + 
  labs( x = "Date", y = "Cumulative cases", title = "Covid-19 Cases in USA" )

usa_traffic <- ggplot( data = usa_air_traffic, aes( x = Date, y = Cumulative_flights ) ) + 
  geom_line( stat = "identity", color = "blue" ) + 
  labs( x = "Date", y = "Cumulative flights", title = "Total flights in USA" )

grid.arrange( usa_cases, usa_traffic, nrow = 1 )

```

checking if covid spread or cases increased in summer season and festival season due to more travel and family gatherings

```{r}

# Get the library.
library( plotrix )
#library(scales)

# summer holidays in USA mid may to about mid August
 
usa_covid$year <- format( usa_covid$Date_reported, "%y" )              # Extract year

usa_covid_20 <- usa_covid[which(usa_covid$year == '20'),]

usa_covid_21 <- usa_covid[ which( usa_covid$year == '21' ), ]

usa_covid_22 <- usa_covid[ which( usa_covid$year == '22' ), ]



usa_covid_20$month <- format( usa_covid_20$Date_reported, "%m" )              # Extract month

# grouping by month and adding the covid cases to get the total cases registered in that whole month
usa_covid_sum_20 <- usa_covid_20 %>% group_by( month ) %>% summarise( usa_sum_case_20 = sum( New_cases ) )

# plotting bar graphs by passing in month as x and cases as y
# setting each bar width to 1 and writing the value in black
ggplot( usa_covid_sum_20, aes( x = month, y = usa_sum_case_20, fill = month, width = 1)) + geom_bar( stat = "identity" ) + geom_text( aes(label = usa_sum_case_20 ), vjust = 2, colour = "black")




usa_covid_21$month <- format( usa_covid_21$Date_reported, "%m" )              # Extract month

# grouping by month and adding the covid cases to get the total cases registered in that whole month
usa_covid_sum_21 <- usa_covid_21 %>% group_by( month ) %>% summarise( usa_sum_case_21 = sum( New_cases ) )


# plotting bar graphs by passing in month as x and cases as y
# setting each bar width to 1 and writing the value in black
ggplot( usa_covid_sum_21, aes( x = month, y = usa_sum_case_21, fill = month, width = 1 )) + geom_bar( stat = "identity" ) + geom_text( aes( label = usa_sum_case_21 ), vjust = 2, colour = "black" )




usa_covid_22$month <- format( usa_covid_22$Date_reported, "%m" )              # Extract month

# grouping by month and adding the covid cases to get the total cases registered in that whole month
usa_covid_sum_22 <- usa_covid_22 %>% group_by( month ) %>% summarise( usa_sum_case_22 = sum( New_cases ) )


# plotting bar graphs by passing in month as x and cases as y
# setting each bar width to 1 and writing the value in black
ggplot( usa_covid_sum_22, aes( x = month, y = usa_sum_case_22, fill = month, width = 1 )) + geom_bar( stat = "identity" ) + geom_text( aes( label = usa_sum_case_22 ), vjust = 2, colour = "black" )


# ------------------------------------------------------------------------------------------------------------------



# summer holidays in India Mid may to July 1st

india_covid$year <- format( india_covid$Date_reported, "%y" )              # Extract year

india_covid_20 <- india_covid[which(india_covid$year == '20'),]

india_covid_21 <- india_covid[ which( india_covid$year == '21' ), ]

india_covid_22 <- india_covid[ which( india_covid$year == '22' ), ]




india_covid_20$month <- format( india_covid_20$Date_reported, "%m" )              # Extract month

# grouping by month and adding the covid cases to get the total cases registered in that whole month
india_covid_sum_20 <- india_covid_20 %>% group_by( month ) %>% summarise( india_sum_case_20 = sum( New_cases ) )


# plotting bar graphs by passing in month as x and cases as y
# setting each bar width to 1 and writing the value in black
ggplot( india_covid_sum_20, aes( x = month, y = india_sum_case_20, fill = month, width = 1 )) + geom_bar( stat = "identity" ) + geom_text( aes( label = india_sum_case_20 ), vjust = 2, colour = "black" )





india_covid_21$month <- format( india_covid_21$Date_reported, "%m" )              # Extract month

# grouping by month and adding the covid cases to get the total cases registered in that whole month
india_covid_sum_21 <- india_covid_21 %>% group_by( month ) %>% summarise( india_sum_case_21 = sum( New_cases ) )


# plotting bar graphs by passing in month as x and cases as y
# setting each bar width to 1 and writing the value in black
ggplot( india_covid_sum_21, aes( x = month, y = india_sum_case_21, fill = month, width = 1 )) + geom_bar( stat = "identity" ) + geom_text( aes( label = india_sum_case_21 ), vjust = 2, colour = "black" )





india_covid_22$month <- format( india_covid_22$Date_reported, "%m" )              # Extract month

# grouping by month and adding the covid cases to get the total cases registered in that whole month
india_covid_sum_22 <- india_covid_22 %>% group_by( month ) %>% summarise( india_sum_case_22 = sum( New_cases ) )


# plotting bar graphs by passing in month as x and cases as y
# setting each bar width to 1 and writing the value in black
ggplot( india_covid_sum_22, aes( x = month, y = india_sum_case_22, fill = month, width = 1 )) + geom_bar( stat = "identity" ) + geom_text( aes( label = india_sum_case_22 ), vjust = 2, colour = "black" )



# --------------------------------------------------------------------------------------------------------------------




germany_covid$year <- format( germany_covid$Date_reported, "%y" )              # Extract year

germany_covid_20 <- germany_covid[ which( germany_covid$year == '20' ),]

germany_covid_21 <- germany_covid[ which( germany_covid$year == '21' ), ]

germany_covid_22 <- germany_covid[ which( germany_covid$year == '22' ), ]




germany_covid_20$month <- format( germany_covid_20$Date_reported, "%m" )              # Extract month

# grouping by month and adding the covid cases to get the total cases registered in that whole month
germany_covid_sum_20 <- germany_covid_20 %>% group_by( month ) %>% summarise( germany_sum_case_20 = sum( New_cases ) )


# plotting bar graphs by passing in month as x and cases as y
# setting each bar width to 1 and writing the value in black
ggplot( germany_covid_sum_20, aes( x = month, y = germany_sum_case_20, fill = month, width = 1 )) + geom_bar( stat = "identity" ) + geom_text( aes( label = germany_sum_case_20 ), vjust = 2, colour = "black" )




germany_covid_21$month <- format( germany_covid_21$Date_reported, "%m" )              # Extract month

# grouping by month and adding the covid cases to get the total cases registered in that whole month
germany_covid_sum_21 <- germany_covid_21 %>% group_by( month ) %>% summarise( germany_sum_case_21 = sum( New_cases ) )


# plotting bar graphs by passing in month as x and cases as y
# setting each bar width to 1 and writing the value in black
ggplot( germany_covid_sum_21, aes( x = month, y = germany_sum_case_21, fill = month, width = 1 )) + geom_bar( stat = "identity" ) + geom_text( aes( label = germany_sum_case_21 ), vjust = 2, colour = "black" )


germany_covid_22$month <- format( germany_covid_22$Date_reported, "%m" )              # Extract month

# grouping by month and adding the covid cases to get the total cases registered in that whole month
germany_covid_sum_22 <- germany_covid_22 %>% group_by( month ) %>% summarise( germany_sum_case_22 = sum( New_cases ) )


# plotting bar graphs by passing in month as x and cases as y
# setting each bar width to 1 and writing the value in black
ggplot( germany_covid_sum_22, aes( x = month, y = germany_sum_case_22, fill = month, width = 1 )) + geom_bar( stat = "identity" ) + geom_text( aes( label = germany_sum_case_22 ), vjust = 2, colour = "black" )


```


Did the total number of vaccines affect the total amount of deaths in a country?

```{r}


# grouping by month and adding the covid cases to get the total cases registered in that whole month
usa_covid_death_sum_20 <- usa_covid_20 %>% group_by( month ) %>% summarise( usa_death_sum_case_20 = sum( New_deaths ) )

# grouping by month and adding the covid cases to get the total cases registered in that whole month
usa_covid_death_sum_21 <- usa_covid_21 %>% group_by( month ) %>% summarise( usa_death_sum_case_21 = sum( New_deaths ) )

# grouping by month and adding the covid cases to get the total cases registered in that whole month
usa_covid_death_sum_22 <- usa_covid_22 %>% group_by( month ) %>% summarise( usa_death_sum_case_22 = sum( New_deaths ) )


# plotting line graphs by passing in month as x and deaths as y
# setting each bar width to 1 and drawing lines in red for this country
usa_20 <- ggplot( data = usa_covid_death_sum_20, aes( x = month, y = usa_death_sum_case_20, group = 1 )) + geom_line( color = "red", size = 1.5 ) + geom_point()
usa_21 <- ggplot( data = usa_covid_death_sum_21, aes( x = month, y = usa_death_sum_case_21, group = 1 )) + geom_line( color = "red", size = 1.5 ) + geom_point()
usa_22 <- ggplot( data = usa_covid_death_sum_22, aes( x = month, y = usa_death_sum_case_22, group = 1 )) + geom_line( color = "red", size = 1.5 ) + geom_point()

grid.arrange( usa_20, usa_21, usa_22, nrow = 1 )



# grouping by month and adding the covid cases to get the total cases registered in that whole month
india_covid_death_sum_20 <- india_covid_20 %>% group_by( month ) %>% summarise( india_death_sum_case_20 = sum( New_deaths ) )

# grouping by month and adding the covid cases to get the total cases registered in that whole month
india_covid_death_sum_21 <- india_covid_21 %>% group_by( month ) %>% summarise( india_death_sum_case_21 = sum( New_deaths ) )

# grouping by month and adding the covid cases to get the total cases registered in that whole month
india_covid_death_sum_22 <- india_covid_22 %>% group_by( month ) %>% summarise( india_death_sum_case_22 = sum( New_deaths ) )


# plotting line graphs by passing in month as x and deaths as y
# setting each bar width to 1 and drawing lines in blue for this country
india_20 <- ggplot( data = india_covid_death_sum_20, aes( x = month, y = india_death_sum_case_20, group = 1 )) + geom_line( color = "blue", size = 1.5 ) + geom_point()
india_21 <- ggplot( data = india_covid_death_sum_21, aes( x = month, y = india_death_sum_case_21, group = 1 )) + geom_line( color = "blue", size = 1.5 ) + geom_point()
india_22 <- ggplot( data = india_covid_death_sum_22, aes( x = month, y = india_death_sum_case_22, group = 1 )) + geom_line( color = "blue", size = 1.5 ) + geom_point()

grid.arrange( india_20, india_21, india_22, nrow = 1 )


# grouping by month and adding the covid cases to get the total cases registered in that whole month
germany_covid_death_sum_20 <- germany_covid_20 %>% group_by( month ) %>% summarise( germany_death_sum_case_20 = sum( New_deaths ) )

# grouping by month and adding the covid cases to get the total cases registered in that whole month
germany_covid_death_sum_21 <- germany_covid_21 %>% group_by( month ) %>% summarise( germany_death_sum_case_21 = sum( New_deaths ) )

# grouping by month and adding the covid cases to get the total cases registered in that whole month
germany_covid_death_sum_22 <- germany_covid_22 %>% group_by( month ) %>% summarise( germany_death_sum_case_22 = sum( New_deaths ) )

# plotting line graphs by passing in month as x and deaths as y
# setting each bar width to 1 and drawing lines in green for this country
germany_20 <- ggplot( data = germany_covid_death_sum_20, aes( x = month, y = germany_death_sum_case_20, group = 1 )) + geom_line( color = "green", size = 1.5 ) + geom_point()
germany_21 <- ggplot( data = germany_covid_death_sum_21, aes( x = month, y = germany_death_sum_case_21, group = 1 )) + geom_line( color = "green", size = 1.5 ) + geom_point()
germany_22 <- ggplot( data = germany_covid_death_sum_22, aes( x = month, y = germany_death_sum_case_22, group = 1 )) + geom_line( color = "green", size = 1.5 ) + geom_point()

grid.arrange( germany_20, germany_21, germany_22, nrow = 1 )

```